### 2023-03-27 uvp6_part.qmd
Intro

Inside the ocean, a cohort of diverse organisms known as zooplankton thrive. Those tiny beasts are so important for our oceans [@Steinberg2017-ri] that we decided to take pictures of them [@Picheral2010-to; @Gorsky2010-ca; @Cowen2008-wq; @Benfield2007-lt; @Lombard2019-ld]. Nonetheless, the amount of images taken by those imaging instruments has become so huge that it is now a massive work to classify them by hand. Fortunately, powerful machine learning techniques have been implemented to tackle the issue of automatic plankton image classification [@Luo2018-oo].

However, it remains a fierce challenge. Within a single taxon, one can find organisms with very different sizes and shapes, organisms with significant morphological differences acquired during their ontogenetic development, partial organisms (e.g. siphonophores) or organisms with various orientations, hindering the automatic classification process [@Benfield2007-lt]. In addition to this global heterogeneity, the great majority of images are not composed of zooplankton. Instead, non-living particles (mostly marine snow) steal the show, producing a highly imbalanced data set towards the latter.

The discussion around automatic plankton classification is a key topic in oceanography and the development of convolutional neural networks (CNNs) within that frame has already lead to promising results [@Luo2018-oo; @Cui2018-nt; @Dai2017-ex; @Campbell2020-cq; @Schroder2019-um; @Bochinski2019-af; @Lee2016-jl; @Cheng2019-ar; @Guo2021-xi; @Ellen2019-nh; @Orenstein2017-ya]. Nevertheless, real-time image processing done at sea is far less mentioned to our knowledge. Usually, images are brought back to the lab and classified with CNNs which, in comparison with features-based classifiers, that is machine learning algorithms using a set of features extracted from the image (e.g. area, eccentricity, gray levels) have outperformed the latter for plankton recognition [@Ellen2019-nh]. Though, it comes at a cost. CNNs are energy consuming hence their implementation for embedded applications can be limited by power consumption [@Biswas2018-uw; @Khabbazan2019-oi; @Oh2017-kj].

This is of particular importance for autonomous underwater vehicles (AUVs) and especially for a new version of Biogeochemical-Argo (BGC-Argo) floats that will be equipped with the Underwater Vision Profiler 6 (UVP6), the portable version of the UVP5 [@Picheral2010-to] to better understand the carbon cycle [@Xing2018-zo]. The lifetime of floats is limited by their batteries hence by the power consumed by its sensors. Therefore, using an embedded CNN to classify plankton images or sending images through iridium transmission to be classified in the lab would not be possible for economical (cost of iridium transmission) and technical reasons. Typically, a BGC-Argo float samples the water column between the surface and 1,000 m, it would therefore have to stay at the surface for a time long enough to transmit all images collected after each profile, consuming energy proportionally to the size of the image set while being subject to biofouling or unpredictable hazards. For all these reasons, it was decided to use a features-based classifier for the development of the embedded plankton recognition algorithm on board the UVP6.

In the literature, such algorithms (e.g. Decision Tree (DT), Random Forest (RF), Support Vector Machine (SVM)) have already been used for plankton recognition (see Table 1 of @Gonzalez2017-tr for an exhaustive list). For instance, @Blaschko2005-ep obtained an accuracy of 71% using a SVM on a set of 982 images among 13 classes using a 10-fold cross-validation (CV), @Sosik2007-cj obtained a global accuracy of 88% using a SVM on a test set of 3300 images covering 22 classes, @Bell2008-yo obtained 82% of accuracy using a 10-fold CV RF algorithm on 53 classes with 10 to 30 images per class. However, it should be noted that this accuracy dropped to 63% after the removal of non-biological classes such as detritus and fiber. For bigger training sets, which will be the case in this study, @Orenstein2017-ya trained a RF model on approximately 650.000 images across 70 classes for an accuracy of 91% on the test set (2.75 million images). Though, one class in that test set contained 1.6 million images and if the classifier had put every object in that class, the accuracy would have been around 60% so caution must be applied when interpreting performance metrics. @Faillettaz2016-wr trained a RF model on 6000 images across 14 classes, reaching an accuracy of 56% on a test set of 1.5 million objects, discarding those below a certain probability threshold. This allowed to improve the global accuracy by 16% but resulted in the loss of 72% of objects in the predicted data set.

It is quite difficult to compare all methods due to the diversity of imaging instruments hence the variety of image formats (e.g. resolution), the number of classes (trade off between the model performance and the number of classes as illustrated in @Fernandes2009-hw, the training set (e.g. number of images per class, diversity, balance), the scientific objective (e.g. precision-recall trade off), the type and number of features and the algorithm used. In many cases, RF seemed to be the best classifier [@Gorsky2010-ca; @Faillettaz2016-wr] but in embedded applications, energy constraints need to be taken into account. Therefore, the model needs to be small and the classification should be easy and fast. Gradient Tree Boosting (GTB) seems to offer this compromise when compared to RF (model size) and SVM (computation complexity). For these reasons, we chose XGBoost [@Chen2016-xt], an optimized implementation of GTB, also known to be one of the best assets to win Kaggle's machine learning competitions.

In this paper, we present the results of an embedded zooplankton recognition algorithm using XGBoost on UVP5 data. We show how we built the embedded recognition (ER) model under strict technical constraints, what were our compromises and what techniques we used to improve the classification. We also compare the performances of our features-based algorithm with a state-of-the-art neural network as if energy was not a limiting factor. Finally, we apply the method and follow the recommendations learnt from the preliminary UVP5 model to the UVP6 data set to build the ER model that will be included in the UVP6 for AUVs.

@@

GTB generates a sequence of trees, each one trying to cope with the errors made by their predecessors. XGBoost is designed to be an optimized and regularized version of GTB. To learn from the data and derive a model, XGBoost first minimizes a regularized objective function $L$

>>>>

Inside the ocean, zooplankton constitutes a cohort of very diverse organisms. They are important enough for the functioning of our oceans [@Steinberg2017-ri] to motivate the development of a large number of instruments dedicated to taking pictures of them [@Picheral2010-to; @Gorsky2010-ca; @Cowen2008-wq; @Benfield2007-lt; @Lombard2019-ld]. The amount of images taken by those instruments has quickly become so large that it is now unfeasible to classify them manually, through the observation of human operators. Fortunately, powerful machine learning techniques have been implemented to tackle the issue of automatic plankton image classification [@Luo2018-oo].

However, it remains a fierce challenge. Within a single taxon, individuals with very different sizes and shapes coexist, due to significant morphological changes during ontogeny, imaging of portions of organisms only (e.g. siphonophores) or organisms taken from various angles. This variability hinders the performance of automatic classification [@Benfield2007-lt]. In addition to this global heterogeneity, for in situ instruments, the large majority of images are not of zooplankton: non-living particles (mostly marine snow) largely dominate, resulting in highly imbalanced data sets.

Automatic plankton classification is an active research topic. Within this context, the application of Convolutional Neural Networks (CNNs) has already lead to promising results [@Luo2018-oo; @Cui2018-nt; @Dai2017-ex; @Campbell2020-cq; @Schroder2019-um; @Bochinski2019-af; @Lee2016-jl; @Cheng2019-ar; @Guo2021-xi; @Ellen2019-nh; @Orenstein2017-ya]. Indeed, CNNs have outperformed the previous features-based classifiers (i.e. machine learning algorithms using a set of features extracted from the image such as area, eccentricity, gray levels, etc.) [@Ellen2019-nh]. Nevertheless, real-time image processing, performed at sea, in the instrument, has not been tackled, to our knowledge. Usually, images are brought back to the lab and classified with classic machine learning classifiers and, increasingly, CNNs. But CNNs are energy consuming, even just for inference (i.e. processing and classification of new images, without retraining), and their implementation for embedded applications can be limited by power consumption [@Biswas2018-uw; @Khabbazan2019-oi; @Oh2017-kj].

This is of particular importance for autonomous underwater vehicles (AUVs) and especially for a new version of Biogeochemical-Argo (BGC-Argo) floats that will be equipped with the Underwater Vision Profiler 6 (UVP6), the portable version of the UVP5 [@Picheral2010-to] to better understand the carbon cycle [@Xing2018-zo]. The lifetime of floats is limited by their batteries hence by the power consumed by its sensors. The floats are not recovered and their data has to be sent through satellite communication. Therefore, using an embedded CNN to classify plankton images onboard or sending images through the satellite link are both impossible with currently available hardware. [JOI: ici il manque une justif de pq c'est poas possible de mettre un CNN dans le flotteur; il y a qq chiffres dans l'étude de FOTONOWER; tu peux aussi parler des cartes dédiées mais qui sont encore largement trop consommatrices => il manque un chiffre sur la conso énergétique en W attendue d'un capteur; un ordre de grandeur suffit]. Typically, BGC-Argo floats sample the water column between 1,000 m [JOI: c'est pas 2000?] and the surface; in this range, the UVP6 is expected to take images of hundreds of planktonic objects and, given the bandwidth of the current connection, it would require the float to stay [JOI: n minutes/hours/days] to transmit them [JOI: pareil, on avait des stats et même des estimations de temps, il faut donner les chiffres], consuming energy proportionally to the amount of data sent and being subject to biofouling or unpredictable hazards. For these reasons, it was decided to use a simpler features-based classifier for the development of the embedded plankton recognition algorithm on board the UVP6.

Such algorithms (e.g. Decision Tree (DT), Random Forest (RF), Support Vector Machine (SVM)) have already been used for plankton classification (see Table 1 of @Gonzalez2017-tr for an exhaustive list at that time). For instance, @Blaschko2005-ep obtained an accuracy of 71% using a SVM on a set of 982 images among 13 classes using a 10-fold cross-validation (CV), @Sosik2007-cj obtained a global accuracy of 88% using a SVM on a test set of 3300 images covering 22 classes, @Bell2008-yo obtained 82% of accuracy using a 10-fold CV RF algorithm on 53 classes with 10 to 30 images per class. However, it should be noted that this accuracy dropped to 63% after the removal of non-biological classes such as detritus and fiber. Using a larger data set or 650.000 images, similar to the one used in this study, @Orenstein2017-ya trained a RF model over 70 classes that yielded an accuracy of 91% on the test set (2.75 million images). Though, one class in that test set contained 1.6 million images and a dummy classifier that would have put every object in that class would have given an accuracy of about 60%; so caution must be applied when interpreting this global performance metric. @Faillettaz2016-wr trained a RF model on 6000 images across 14 classes, reaching an accuracy of 56% on a test set of 1.5 million objects, discarding those below a certain probability threshold. This allowed to improve the global accuracy by 16% but resulted in the loss of 72% of objects in the predicted data set. [JOI: Il y a une review par Irisson et al qui fait un petit bilan de tout ça ;-)]

It is quite difficult to compare all methods due to the diversity of imaging instruments resulting in a variety of image resolutions, numbers of classes (while there is a trade off between model performance and the number of classes as illustrated in @Fernandes2009-hw), training sets (e.g. number of images per class, diversity, balance), scientific objectives (e.g. precision-recall trade off), types and numbers of features used and classification algorithms. In many cases, RF seemed to be the best classifier [e.g. @Gorsky2010-ca] [JOI: je crois que c'est plutôt grosjean]. For embedded applications, the number of parameters in the model need to be small for its memory footprint to be reduced and the computation of new classifications should be fast. Gradient Boosted Trees (GBT) seem to offer this compromise when compared to RF (model size) and SVM (computation complexity). For these reasons, we chose to classify images using the XGBoost library [@Chen2016-xt], an optimized implementation of GBT, known to be high performing in machine learning competitions (on the Kaggle website for example).

In this paper, we present the results of an embedded zooplankton classification algorithm using XGBoost on UVP5 data. We show how we built the embedded classification (EC) [JOI: UVPec = UVP embedded classification => soit cohérent; et puis recognition c'est pas trop ça en Anglais] model under strict technical constraints, what were our compromises and what techniques we used to improve the classification. We also compare the performances of our features-based algorithm with a state-of-the-art neural network as if energy was not a limiting factor. Finally, we apply the method and follow the recommendations learnt from the preliminary UVP5 model to the UVP6 data set to build the EC model that will be included in the UVP6 for AUVs.

@@

GBT generates a sequence of trees, each one trying to cope with the errors made by their predecessors. XGBoost is designed to be an optimized and regularized version of GBT. To learn from the data and derive a model, XGBoost first minimizes a regularized objective function $L$
